---
title: "Day 7 Session 1"
format: html
editor: visual
---

```{r}
library(palmerpenguins)
library(tidyverse)

```

### Data wrangling refreshser

1.  starting from penguins, write a single piped sequence in which you only include penguins at dream and biscoe islands, remove the year and sex variables, add a new column body_mass_kg coverted from g to kg, rename the island variable to location.

```{r}

penguins_1 <- penguins %>% 
  filter(island %in% c("Dream","Biscoe")) %>% 
  dplyr::select(-year, -sex) %>% 
  mutate(body_mass_kg = body_mass_g / 1000) %>% 
  rename(location = island)


```

2.  Starting from penguins write a single piped sequence to :

-   limit to adelie penguins

-   remove any observations to where flipper_length_mm is NA (drop_na)

-   group data by sex

-   make a summary table containing the mean, sd, and sample size of flipper lengths for male and female adelie penguins

```{r}
adelie_penguins <- penguins %>% 
  filter(species == "Adelie") %>% 
  drop_na(flipper_length_mm, sex) %>% 
  group_by(sex) %>% 
  summarize(mean_flipper_length = mean(flipper_length_mm),
            sd_flipper_length = sd(flipper_length_mm),
            sample_size = n()) 
  

  
```

```{r}
penguins %>% 
  group_by(species, island, sex) %>% 
  summarize(sample_size = n())

#count is just wrapping together group_by +summarizE()+n() to get counts of observations. 
```

```{r}
animal <- data.frame(
  stringsAsFactors = FALSE,
          location = c("lagoon", "bluff", "creek", "oaks", "bluff"),
                      species = c("bobcat",
                                  "coyote","fox","squirrel","bobcat"),
                     maturity = c("adult",
                                  "juvenile","adult","juvenile","adult")
           )
  
site <- data.frame(
  stringsAsFactors = FALSE,
          location = c("beach", "lagoon", "bluff", "oaks"),
            full_site_name = c("Goleta Beach",
                               "UCSB Lagoon","Ellwood Mesa",
                               "Fremont Campground"),
      jurisdiction = c("SB City", "UCSB", "SB City", "USFS")
        )
```

```{r}
full_ex <- full_join(animal, site)
```

```{r}
inner_ex <- inner_join(animal,site)
```

```{r}
left_join_ex <- left_join(animal, site)
```
